---
title: <p align="center"> <b> Assessment 1 </b> </p>
output: html_document
---
<p align="center"> *Author: PG* </p>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='figure/')
setwd("C:/Users/Paula/Documents/Coursera/5-ReproducibleResearch/Assessment_1/RepData_PeerAssessment1/")
library(knitr)
```

##**Loading and preprocesing the data**
```{r Loading and preprocesing data}
data<- read.csv("activity.csv") #Reading the data
head(data, 3) #Exploring the data
good<- complete.cases(data) #Removing rows with missing values
data<- data[!good==FALSE, ]
head(data, 3) #Checking
```

##**What is mean total number of steps taken per day?**
Total number of steps taken per day
```{r Total steps per day}
totalsteps<- aggregate(data$steps, list(Date=data$date), sum)
head(totalsteps, 3)
```

Histogram of the total number of steps taken each day
```{r Histogram}
barplot(totalsteps$x, horiz = TRUE, col="aquamarine", main="Total steps taken each day", xlab="Steps", ylab="Day", axes = F)
axis(1, las=3, cex.axis=0.75)
axis(2, las=1, cex.axis=0.75)
```

Mean and Median of the total number of steps taken per day
```{r Mean and Median}
mean(totalsteps$x)
median(totalsteps$x)
```

##**What is the average daily activity pattern?**
Time series plot of the 5-minute interval and the average number of steps taken
```{r Time series plot}
meansteps<- aggregate(data$steps, list(Interval=data$interval), mean) #Calculating mean steps for each interval
plot(meansteps$Interval, meansteps$x, type="l", main="Time series plot", xlab="5-minute interval", ylab="Average number of steps", col="aquamarine", lwd=4)
```

Calculating 5-minute interval, on average across all the days, that contains the maximum number of steps
```{r Maximum number of steps}
meansteps$Interval[max(meansteps$x)]
```


##**Imputing missing values**
Total number of missing values in the dataset
```{r Missing values}
data<- read.csv("activity.csv") #Reading the data again (I eliminated NAs rows in the firsts steps)
length(which(is.na(data)))
```

Filling missing values with the mean for that 5-minute interval and creating a new dataset with the missing data filled
```{r Filling missing values}
nonas<- data
nonas$steps[is.na(nonas$steps)]<- mean(nonas$step, na.rm=T)
head(nonas, 3)
```

Histogram of the total number of steps taken each day
```{r Histogram of steps}
totalsteps2<- aggregate(nonas$steps, list(Date=nonas$date), sum)
barplot(totalsteps2$x, horiz = TRUE, col="aquamarine", main="Total steps taken each day- New dataset", xlab="Steps", ylab="Day", axes = F)
axis(1, las=3, cex.axis=0.75)
axis(2, las=1, cex.axis=0.75)
```

Mean and Median of total number of steps taken per day
```{r Mean and Median newdataset}
mean(totalsteps2$x)
median(totalsteps2$x)
```
Like I had removed already the misssing values, the impact of imputing missing data on the estimates of the total daily number of steps is indiscernible.


##**Are there differences in activity patterns between weekdays and weekends?**
Creating a new factor variable in the dataset with two levels indicating whether a given date is a weekday or weekend day
```{r Creating new factor}
total<- weekdays(as.POSIXlt(data$date), abbreviate= TRUE)
data$day<- as.factor(total)
levels(data$day)
library(car)
data$day<- recode(data$day, "c('ju.', 'lu.', 'ma.', 'mi.', 'vi.')='weekday'; c('sá.', 'do.')='weekend'")
levels(data$day)
```

Panel plot containing a time series plot of the 5-minutes interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r Panel plot}
library(lattice)
attach(data)
xyplot(steps~interval|factor(day), data=data, main="Time series plot", ylab="Number of steps", xlab="Interval", layout=c(1,2), type="l", col="aquamarine")
```
